### 2023-06-19 paper/FPF-WG1.tex
abstract

{\it \small
    ~$^1$ UC Irvine  \\[0.1cm]
    ~$^2$Department of Physics and Astronomy, Vrije Universiteit, NL-1081 HV Amsterdam\\[0.1cm]
    ~$^3$Nikhef Theory Group, Science Park 105, 1098 XG Amsterdam, The Netherlands\\[0.1cm]
    ~$^4$National Centre for Nuclear Research, Pasteura 7, Warsaw, PL-02-093, Poland \\[0.1cm]
 }

@@

Proton-proton collisions at the LHC produce an intense collimated flux of neutrinos
characterised by the largest energies (up to several TeV) achieved in laboratory experiments.
%
Until recently, these neutrinos escaped detection and hence were lost for physics.
%
The observation in March 2023 of LHC neutrinos by the FASER$\nu$ and SND@LHC experiments
heralds the beginning of a ``LHC neutrino era'' offering unprecedented scientific opportunities
for particle and astroparticle physics both within the Standard Model and beyond it.
%
Here we assess the expected impact of LHC neutrino deep-inelastic scattering (DIS) structure functions
on the quark and gluon substructure of nucleons and heavy nuclei.
%
In addition to projections for FASER$\nu$ and SND@LHC based on the full Run III luminosity,
we also consider the experiments entering the proposed
 Forward Physics Facility (FPF) to operate concurrently with the HL-LHC.
%
 We generate pseudo-data for inclusive and charm DIS structure functions for these various experiments,
model their  systematic uncertainties, and quantify their impact
on proton and nuclear PDFs by means of both  Hessian profiling within {\sc\small xFitter}
and using the NNPDF framework.
%
We find the LHC neutrino scattering, specially
for the high-statistics  experiments at the FPF, provides  stringent constraints
concerning the separation between quark and antiquark PDFs of different flavours.
%
Our analysis demonstrates that detection and analysis of LHC far-forward neutrinos  effectively
extends the LHC with a charged-current counterpart of the Electron-Ion Collider (EIC)
with comparable kinematic reach and hence realises a ``Neutrino-Ion Collider'' at CERN.

>>>>

{\it \small
    ~$^1$ Department of Physics and Astronomy, University of California, Irvine, CA 92697 USA  \\[0.1cm]
    ~$^2$Department of Physics and Astronomy, Vrije Universiteit, NL-1081 HV Amsterdam\\[0.1cm]
    ~$^3$Nikhef Theory Group, Science Park 105, 1098 XG Amsterdam, The Netherlands\\[0.1cm]
    ~$^4$National Centre for Nuclear Research, Pasteura 7, Warsaw, PL-02-093, Poland \\[0.1cm]
 }

@@

Proton-proton collisions at the LHC produce an intense collimated beam of high-energy neutrinos
in the forward direction.
%
Their recent  observation by the FASER$\nu$ and SND@LHC experiments
indicates that this hitherto discarded beam becomes now available for scientific studies.
%
Neutrino deep-inelastic scattering (DIS) is known
to provide a sensitive probe of the quark flavour separation in nucleons
and nuclei, in particular strangeness.
%
Here we assess the expected impact of DIS neutrino structure functions at the LHC
on the parton distribution functions (PDFs) of protons and heavy nuclei.
%
We consider projections for  FASER$\nu$ and SND@LHC based on the Run III luminosity
and for the  experiments hosted by the
 Forward Physics Facility (FPF) operating concurrently with the HL-LHC.
 %
 By means of both Hessian profiling within {\sc\small xFitter} and
 of the  NNPDF open-source framework,
 we demonstrate the significant PDF impact of the high-statistics FPF data
 on the region $10^{-3}\lsim x \lsim 0.7$, comparable to that covered by the
 Electron-Ion Collider (EIC) with neutral-current scattering.
%
 Our analysis highlights how exploiting the potential
 of this unique neutrino beam  effectively
extends the LHC with a ``Neutrino-Ion Collider''.
